Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-06 paper/paper.tex
Clojure-independent implementation of sample and observe

\begin{figure}
	\centering
    \includegraphics[trim={12pt 0 0 0},scale=0.7]{same-customer-time.pdf}
    \caption{Arrival time distribution for a single customer.}
    \label{fig:dist-same}
\end{figure}

\begin{figure}
	\centering
    \includegraphics[trim={12pt 0 0 0},scale=0.7]{different-customers-times.pdf}
    \caption{Arrival time distributions for two customers.}
    \label{fig:dist-diff}
\end{figure}

@@

The purpose of probabilistic forms \texttt{sample} and
\texttt{observe} is to interrupt deterministic computation and
transfer control to the inference algorithm.  Practically, this
is achieved through returning \textbf{checkpoints} --- Clojure
values of the corresponding types
(\texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}sample} or
\texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}observe}).
The types contain fields specific to each form, as well as the
continuation; calling the continuation resumes the computation.
Checkpoints expose the program state to the inference algorithm,
and the updated state is re-injected into the computation when
the continuation is called:
\begin{lstlisting}[style=default]
=> (cps-of-expression '(sample dist) 'cont)
(->sample dist cont $state)
=> (cps-of-expression '(observe dist v) 'cont)
(->observe dist v cont $state)
\end{lstlisting}
Here \texttt{->sample} and \texttt{->observe} in the CPS-transformed
expressions are constructors for Clojure records, and they take 
values of their fields as arguments. How the state is updated, 
the number of times the continuation is called, and the value
passed to the continuation of \texttt{sample} depend on the
inference algorithm executing the program.
Section~\ref{sec:inference} provides more detail on internals of
inference algorithms and their interaction with probabilistic
programs.

>>>>

\begin{figure}
	\centering
    \includegraphics[trim={12pt 0 0 0},scale=0.6]{same-customer-time.pdf}
    \caption{Arrival time distribution for a single customer.}
    \label{fig:dist-same}
\end{figure}

\begin{figure}
	\centering
    \includegraphics[trim={12pt 0 0 0},scale=0.6]{different-customers-times.pdf}
    \caption{Arrival time distributions for two customers.}
    \label{fig:dist-diff}
\end{figure}

@@

The purpose of probabilistic forms \texttt{sample} and
\texttt{observe} is to interrupt deterministic computation and
transfer control to the inference algorithm.  Practically, this
is achieved by  returning a \textbf{checkpoint} instead of
invoking a continuation in a CPS-transformed program.  In
Clojure, checkpoints are represented by values of the
corresponding record types
(\texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}sample} or
\texttt{anglican.{\linebreak[0]}trap.{\linebreak[0]}observe}).
The types contain fields specific to each form, as well as the
continuation; calling the continuation resumes the computation.
\begin{lstlisting}[style=default]
=> (cps-of-expression '(sample dist) 'cont)
(->sample dist cont $state)
=> (cps-of-expression '(observe dist v) 'cont)
(->observe dist v cont $state)
\end{lstlisting}
Here \texttt{->sample} and \texttt{->observe} in the
CPS-transformed expressions are constructors for Clojure
records, and they take values of their fields as arguments.

Upon encountering a checkpoint, the inference algorithm computes
the updated program state and the value to be passed to the
continuation.  How the state is updated, the number of times the
continuation is called, and the value passed to the continuation
of \texttt{sample} depend on the inference algorithm executing
the program.  Section~\ref{sec:inference} provides more detail
on internals of inference algorithms and their interaction with
probabilistic programs.
